A simulation-based assessment of the relation between Stone Age sites and relative sea-level change along the Norwegian Skagerrak coast

### 2022-01-25 analysis/paper/paper.Rmd
Minor edits
Applying exponential decay to the shoreline dating

Accepting that shoreline dating appears to loose utility around the transition to the Late Neolithic, the results for the second row of Figure \@ref(fig:results) is given again in Figure \@ref(fig:results2), excluding all simulation results younger than 2500 BCE. Furthermore, all negative values have here been set to zero, under the assumption that these result from uncertainty or errors in the data, and not actual site locations. The best point estimate for the distance between site and shoreline for the pre-Late Neolithic is, following from the skewed distribution, given by the median at `r round(median(distances2$vertdist))`m. However, 95% of the distance values fall within the range `r round(quantile(distances2$vertdist, 0))`--`r round(quantile(distances2$vertdist, 0.95))`m. That is, for 95% of the cases, the shoreline was situated on or down to m below the site location. This is constrained further when only considering the Mesolithic dates, with a median offset of `r round(median(distances3$vertdist))`m and with 95% of distance measures within the range `r round(quantile(distances3$vertdist, 0))`--`r round(quantile(distances3$vertdist, 0.95))`m. When applying the method of shoreline dating to Stone Age sites in the region, it could thus be recommended, given no other information, to allow for a potential deviation between site and shoreline from `r round(quantile(distances2$vertdist, 0))`--`r round(quantile(distances2$vertdist, 0.95))` in vertical distance. This can be reduced to `r round(quantile(distances3$vertdist, 0))`--`r round(quantile(distances3$vertdist, 0.95))`m for Mesolithic sites. The sites treated here have been given a shoreline date using these values in Figure. However,

@@

```{r results2, echo = FALSE, fig.cap = "Histograms showing distance from shoreline using only dates older than 2500 BCE that correspond to the site inventory. Negative values have been set to zero.", out.width = "525px", fig.align="center"}
knitr::include_graphics(here("analysis/figures/results2.png"))
```

```{r results3, echo = FALSE, fig.cap = "Histograms showing distance from shoreline using only dates older than 4000 BCE that correspond to the site inventory. Negative values have been set to zero.", out.width = "525px", fig.align="center"}
knitr::include_graphics(here("analysis/figures/results3.png"))
```

```{r shoredate, echo = FALSE, fig.cap = "The result of shoreline dating the analysed sites allowing for a sea-level potentially 4m above to 16m below the lowest site elevation. These are plotted against the oldest radiocarbon-informed dates for each site.", out.width = "525px", fig.align="center"}
knitr::include_graphics(here("analysis/figures/shoredate.png"))
```

>>>>

Accepting that shoreline dating appears to loose utility around the transition to the Late Neolithic, the results for the second row of Figure \@ref(fig:results) is given again in Figure \@ref(fig:results2), excluding all simulation results younger than 2500 BCE. Furthermore, all negative values have here been set to zero, under the assumption that these result from uncertainty or errors in the data, and not actual site locations. The best point estimate for the distance between site and shoreline for the pre-Late Neolithic is, following from the skewed distribution, given by the median at `r round(median(distances2$vertdist))`m. However, 95% of the distance values fall within the range `r round(quantile(distances2$vertdist, 0))`--`r round(quantile(distances2$vertdist, 0.95))`m. That is, for 95% of the cases, the shoreline was situated on or down to m below the site location. This is constrained further when only considering the Mesolithic dates, with a median offset of `r round(median(distances3$vertdist))`m and with 95% of distance measures within the range `r round(quantile(distances3$vertdist, 0))`--`r round(quantile(distances3$vertdist, 0.95))`m. When applying the method of shoreline dating to Stone Age sites in the region, it could thus be recommended, given no other information, to allow for a potential deviation between site and shoreline from `r round(quantile(distances2$vertdist, 0))`--`r round(quantile(distances2$vertdist, 0.95))` in vertical distance. This can be reduced to `r round(quantile(distances3$vertdist, 0))`--`r round(quantile(distances3$vertdist, 0.95))`m for Mesolithic sites. The sites treated here have been given a shoreline date using these values in Figure. However,      
An exponential function has been fit to the distribution. While it is clear that this does not perfectly match the data, it would seem reasonable from a theoretical perspective to assume that a process of exponential decay explains this relationship. The accumulation of distances-values on the 0m mark likely follow from forcing negative values to zero, from the resolution of the spatial data, and from defining intersecting sea- and site polygon as having a distance of zero.

@@

```{r results2, echo = FALSE, fig.cap = "Histograms showing distance from shoreline using dates corresponding to the site inventory. Negative values have been set to zero. The first row only includes dates older than 2500 BCE and the second row only dates older than 4000 BCE.", out.width = "525px", fig.align="center"}
knitr::include_graphics(here("analysis/figures/results2.png"))
```

```{r shoredate_sites, echo = FALSE, fig.cap = "The result of shoreline dating the analysed sites by mean elevation of the site polygon, accounting for distance to closest isobases and employing the exponential decay function identified in Figure 7 for the distance between sites and shoreline. The shoreline dates in gray are plotted against the oldest radiocarbon-informed dates for each site in black. The red line indicates the 95% probability range for the shoreline date.", out.width = "525px", fig.align="center"}

knitr::include_graphics(here("analysis/figures/shoredate.png"))
```
